The tree line ecotone punctuates the abundance and turnover of plants, pollinators, and their interactions

### 2021-06-09 code/ms2_data.Rmd
updated ms2

## Generate a data frame that represents visitation and survey abundance
```{r}
# BB visitation data
visitation <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  filter(!bb.sp %in% c("humi")) %>% # humi too rare
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>% # visits per BB-FL pair
  ungroup() %>%
  left_join(pseudodates) %>% # Now we will replace unmatched dates with their pseudo dates
  replace_na(replace = list(mismatch = 0)) %>% # set mismatch to zero for perfectly aligned dates
  filter(mismatch <= 7) %>% # this gets rid of irreconcilable dates (like the t4 dates mentioned above)
  mutate(date2 = if_else(is.na(pseudo.date), date, pseudo.date)) %>% # replace date with pseudodate for all site-dates lacking perfect alignments
  select(-c(date, pseudo.date, mismatch)) %>%
  rename(date = date2) %>%
  group_by(site, date) %>%
  complete(bb.sp, plant.sp, fill = list(visits = 0)) # generate all possible BB-FL combos for each site date and fill in zeros for visits

# Floral cover data
flower_cover <- read_csv("../data/processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>% # ignore plants that were not blooming
  group_by(site, date, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover))

# Get shared dates after adding pseudodates. This will let us filter out survey dates that don't
# have corresponding visitation sampling dates
shared_dates_with_pseudo <- inner_join(visitation, flower_cover, by = c("site", "date")) %>%
  select(site, date) %>%
  distinct()

# Join visitation and flower cover data

## First we need a list of all BB spp. by site and date
sp_by_site <- visitation %>% 
  ungroup() %>%
  select(site, date, bb.sp) %>%
  distinct()

## Then we need a list of unvisited plant.sp by site and date
unvisited <- flower_cover %>% 
  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
  full_join(visitation) %>% # join to visitation data
  filter(is.na(bb.sp)) %>% # select only the site-date-plants that were totally unvisited (and therefore have no BB spp to join to)
  select(-bb.sp) %>% # drop bb.sp variable because I just want the plants
  left_join(sp_by_site) %>% # join in the species list by site and date, thus populating the unvisited plants with all the BB spp recorded for each site-date
  replace_na(replace = list(visits = 0)) # replace NAs (created in the line above) with zero

## Now we can combine everything into a single data frame
visit_cover <- flower_cover %>% 
  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
  full_join(visitation) %>% # join to visitation data
  replace_na(replace = list(flower.cover = 0.01)) %>% # If a plant was visited but not surveyed, we will assign it the minimum non-zero flower cover value of 0.01.
  filter(!is.na(bb.sp)) %>% # drop the univisited plants
  bind_rows(unvisited) %>% # add them back, with BB spp and zeros
  select(site, date, bb.sp, plant.sp, visits, flower.cover) %>%
  group_by(site, date, bb.sp, plant.sp, flower.cover) %>% # In a few cases, the use of pseudodates causes two sampling period to converge on the same date, which creates apparent duplicate visitation data. We will handle this by averaging the visits.
  summarize(visits = mean(visits)) %>%
  left_join(fl_traits) %>% # add floral traits
  ungroup() %>%
  mutate(k.type.ss = factor(k.type.ss))

## Get total visits by BB sp and site-date; this will be a measure of BB abundance that we can use after we've calculated d'. I could, of course, work this into the biaseval function, but I'll put that off for now.
bb_abundance <- visit_cover %>%
  group_by(site, date, bb.sp) %>%
  summarize(bb.sp.abund = sum(visits)) %>%
  group_by(site, date) %>%
  mutate(bb.sp.prop.abund = bb.sp.abund/sum(bb.sp.abund))

fl_abundance <- visit_cover %>%
  group_by(site, date) %>%
  summarize(total.flower.cover = sum(flower.cover))
```

# 1. Overall bias

## 1.1 Calculate visitation probability per plant morphotype. Instead of keeping the plant species resolved, we will pool visitation by morphotype. The problem with not pooling by morphotype is that more species-rich morphotypes will be penalized simply because their visitation is diluted across more species.
```{r}
prop_visits_ktype <- visit_cover %>%
  group_by(site, date, bb.sp, k.type.ss) %>%
  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
            visits = sum(visits)) %>%
  group_by(site, date) %>%
  mutate(prop.flower.cover = flower.cover/sum(flower.cover))%>% # binary visits
  group_by(site, date, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits),
         bb.sp.abund = sum(visits)) %>% # abundance of each BB sp. per site-date
  ungroup() %>%
  left_join(bb_traits) %>% # add BB traits
  mutate(ktype = factor(k.type.ss), # recode variables
         tongue = factor(pbl.w.class2),
         visits = as.integer(visits)) %>% # so we can run a Poisson model
  left_join(site_data) %>% # add site data
  ungroup() %>%
  mutate(yday = yday(date),
         elev.mean0 = scale(elev.mean),
         yday0 = scale(yday),
         prop.flower.cover0 = scale(prop.flower.cover),
         perc.flower.cover = prop.flower.cover*100,
         bb.sp.abund0 = scale(bb.sp.abund)) %>%
  select(site, elev.mean, date, yday, bb.sp, ktype, visits, 
         prop.visits, flower.cover, perc.flower.cover, bb.sp.abund, tongue,
         elev.mean0, yday0, prop.flower.cover0, bb.sp.abund0) %>%
  mutate(visits.bin = if_else(visits > 0, 1, 0))

write_csv(prop_visits_ktype, "../data/processed_data/prop_visits_ktype.csv")

prop_visits_ktype_pooled <- visit_cover %>%
  group_by(site, date, k.type.ss) %>%
  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
            visits = sum(visits)) %>%
  group_by(site, date) %>%
  mutate(prop.flower.cover = flower.cover/sum(flower.cover),
         prop.visits = visits/sum(visits),
         bb.abund = sum(visits),
         visits = as.integer(visits)) %>%
  mutate(ktype = factor(k.type.ss)) %>%
  left_join(site_data) %>% # add site data
  ungroup() %>%
  mutate(yday = yday(date),
         elev.mean0 = scale(elev.mean),
         yday0 = scale(yday),
         perc.flower.cover = prop.flower.cover*100,
         prop.flower.cover0 = scale(prop.flower.cover),
         bb.abund0 = scale(bb.abund)) %>%
  select(site, elev.mean, date, yday, ktype, visits, 
         prop.visits, bb.abund, perc.flower.cover,
         elev.mean0, yday0, prop.flower.cover0, bb.abund0) %>%
  mutate(visits.bin = if_else(visits > 0, 1, 0),
         year = factor(year(date)))

write_csv(prop_visits_ktype_pooled, "../data/processed_data/prop_visits_ktype_pool.csv")
```

>>>>

## Generate a data frame that represents visitation and survey abundance
```{r}
# BB visitation data
visitation <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  filter(!bb.sp %in% c("humi")) %>% # humi too rare
  left_join(pseudodates) %>% # Now we will replace unmatched dates with their pseudo dates
  replace_na(replace = list(mismatch = 0)) %>% # set mismatch to zero for perfectly aligned dates
  filter(mismatch <= 7) %>% # this gets rid of irreconcilable dates (like the t4 dates mentioned above)
  mutate(date2 = if_else(is.na(pseudo.date), date, pseudo.date)) %>% # replace date with pseudodate for all site-dates lacking perfect alignments
  select(-c(date, pseudo.date, mismatch)) %>%
  rename(date = date2) %>%
  group_by(site, date) %>%
  mutate(visited = rep(1, n())) %>% # binary visitation
  select(site, date, bb.sp, plant.sp, visited) %>%
  left_join(fl_traits) %>%
  select(site, date, bb.sp, plant.sp, visited, plant.genus, plant.family, ktype = k.type.ss) %>%
  unite(sample.id, site, date) %>%
  mutate(sample.id = factor(sample.id),
         bb.sp = factor(bb.sp), 
         ktype = factor(ktype),
         visited = as.integer(visited))

# abundance <- visitation %>%
#   group_by(site, date, bb.sp) %>%
#   summarize(abundance = sum(visited))

# Floral cover data by ktype
flower_cover_ktype <- read_csv("../data/processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>% # ignore plants that were not blooming
  left_join(fl_traits) %>% # add traits
  group_by(site, date, k.type.ss) %>% # sum cover by ktype
  summarize(cover = sum(flower.cover)) %>%
  complete(site, date, k.type.ss, fill = list(cover = 0)) %>% # fill in zeros for absent ktypes
  select(site, date, ktype = k.type.ss, cover) %>%
  unite(sample.id, site, date) %>%
  mutate(sample.id = factor(sample.id),
         ktype = factor(ktype)) %>%
  semi_join(visitation, by = "sample.id")
# 
# # Join cover data to visitation data
# visit_cover_ktype <- visitation %>%
#   left_join(flower_cover_ktype)  %>%
#   replace_na(replace = list(cover = 0.01)) %>% # If a ktype was visited but not surveyed, we will assign it the minimum non-zero flower cover value of 0.01.
#   mutate(individual.id = row_number()) %>%
#   mutate(ktype = case_when(
#     ktype == 0 ~ "null",
#     ktype == 1 ~ "disc",
#     ktype == 2 ~ "funnel",
#     ktype == 3 ~ "bell",
#     ktype == 4 ~ "stalkdisc",
#     ktype == 5 ~ "lip",
#     ktype == 6 ~ "flag",
#     ktype == 7 ~ "head",
#     ktype == 9 ~ "brush"
#   )) %>%
#   unite(sample.id, c(site, date)) %>%
#   mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
#                                           "bell", "stalkdisc", "lip", 
#                                           "flag", "head", "brush")),
#          sample.id = factor(sample.id),
#          bb.sp = factor(bb.sp),
#          visited = as.integer(visited))

# abundance2 <- visit_cover_ktype %>%
#   group_by(site, date, bb.sp) %>%
#   summarize(abundance = sum(visited))
```

```{r}
saveRDS(fl_traits, "../data/processed_data/fl_traits.rds")
saveRDS(visitation, "../data/processed_data/visitation.rds")
saveRDS(flower_cover_ktype, "../data/processed_data/flower_cover_ktype.rds")
#saveRDS(visit_cover_ktype, "../data/processed_data/visit_cover_ktype.rds")
```

### 2021-06-09 ms_1/ms_1.Rmd
updated ms2

## Cache
# opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "../output/")

@@

```{r fig1, echo = FALSE}
knitr::include_graphics("../output/sites_GE.png")
```

@@

```{r fig2, echo = FALSE}
knitr::include_graphics("../output/gam_abund_grid_mod.png")
```

@@

```{r fig3, echo = FALSE}
knitr::include_graphics("../output/beta_stack.png")
```

@@

```{r fig4, echo = FALSE}
knitr::include_graphics("../output/gdm_combo.png")
```

>>>>

## Cache
# opts_chunk$set(cache = 2, cache.path = "ms_2/figures/cache/")

## Plots
opts_chunk$set(fig.path = "../ms_2/figures/")

@@

```{r fig1, echo = FALSE}
knitr::include_graphics("../ms_1/figures/sites_GE.png")
```

@@

```{r fig2, echo = FALSE}
knitr::include_graphics("../ms_1/figures/gam_abund_grid_mod.png")
```

@@

```{r fig3, echo = FALSE}
knitr::include_graphics("../ms_1/figures/beta_stack.png")
```

@@

```{r fig4, echo = FALSE}
knitr::include_graphics("../ms_1/figures/gdm_combo.png")
```
